---
title: "Assignment 05 and Student Survey"
author: "Rapuru, Supraja"
date: '2021-07-25'
output:
  pdf_document: default
  html_document: default
  word_document: default
---
# --- Assignment 05 ---
```{r include = FALSE}
setwd("/Users/Supraja/dsc520")
heights_df <- read.csv("data/r4ds/heights.csv")
tech_spending <- c(18079, 18594, 19753, 20734, 20831, 23029, 23597, 23584, 25525, 27731, 29449)
suicides <- c(5427, 5688, 6198, 6462, 6635, 7336, 7248, 7491, 8161, 8578, 9000)
```
Height Correlations
```{r include = TRUE}
cor(heights_df$height, heights_df$earn)
cor(heights_df$age, heights_df$earn)
cor(heights_df$ed, heights_df$earn)
```
Tech Spending and Suicide Correlation
```{r include= TRUE}
cor(tech_spending, suicides)
```

# --- Student Survey ---
```{r include = FALSE}
library(ggplot2)
library(ggm)
setwd("/Users/Supraja/dsc520")
ss_df <- read.csv("data/student-survey.csv")
```
## Part I
A. Calculate covariance of survey variables  
```{r include= TRUE}
cov(ss_df[, c(1:4)])
```
B. Why use this calculation?  
Covariance in R is calculated by using the cov() function  

C. What do the results indicate?  
1.Time of reading is negatively related to Time of watching TV,
2.Time of reading is negatively related to Happiness.
3.Time of watching TV is positively related to Happiness.
4.As gender is represented as integer, we can ignore the covariance associated with gender.
 

## Part II
A. What measurement is being used for the variables?  
TimeReading: time, in hours (rounded to whole hr)
TimeTV: time, in minutes (rounded to nearest 5 min)
Happiness: looks like percent. It looks like Happiness index varies from 45.67% to 89.52%
Gender - By looking at the values, I assumed that the Gender is measure in boolean.It is not specified that 0 or 1 mean male/female. Need so more info on the variable.

B. Explain what effect changing the measurement being used for the variables   
would have on the covariance calculation.   
Because time is measured in two different ways, out initial covariance calculation  
is not accurate in terms of calculation, but it should still be accurate in terms 
of showing positive or negative covariance.  

C. Would this be a problem? Explain and provide a better alternative if needed.  
Differences in how a type of variable is calculated can be problematic. This is
true not only for time (min v hours) but also for length (in v ft or imperial v metric) 
and any other measurement. We should compare like to like to ensure accuracy.  
  
Here we have the corrected covariance after altering the data so that all time  
is represented in minutes:  
```{r include = FALSE}
TimeReadingMin <- ss_df$TimeReading*60
ss_edit_df <- cbind(ss_df[-c(1)], TimeReadingMin)
ss_edited_df <- ss_edit_df[,c(4,1:3)]
```
```{r include = TRUE}
cov(ss_edited_df[, c(1:4)])
```

## Part III
A. Choose the type of correlation test to perform.  
B. Why this test?  
C. Make a prediction as to whether or not yield +/- correlation.  

checking normality of data

```{r include = FALSE}
ggplot(ss_edited_df, aes(sample=TimeReadingMin)) + stat_qq() + stat_qq_line()
ggplot(ss_edited_df, aes(sample=TimeTV)) + stat_qq() + stat_qq_line()
ggplot(ss_edited_df, aes(sample=Happiness)) + stat_qq() + stat_qq_line()
```
By looking at plots, I can confirm that data is normally distributed.
We can used Perason’s correlation coefficient to check the correlation between variables.

```{r echo = FALSE}
cor.test(ss_edited_df$TimeReadingMin,ss_edited_df$TimeTV)
cor.test(ss_edited_df$TimeReadingMin,ss_edited_df$Happiness)
cor.test(ss_edited_df$Happiness,ss_edited_df$TimeTV)
```

## Part IV

Correlation analysis of:  
A. All variables  
```{r echo = FALSE}
cor(ss_edited_df[, c(1:4)])
```
B. A single correlation between two of the variables  
```{r echo = FALSE}
cor(ss_edited_df$TimeReadingMin, ss_edited_df$TimeTV)
```
C. Repeat above, but set confidence interval at 99%  
```{r echo = FALSE}
cor.test(ss_edited_df$TimeReadingMin, ss_edited_df$TimeTV, conf.level = 0.99)
```
D. Describe what the calculations in the correlation matrix suggest about the relationship between the variables. Be specific.  
The variables TimeReadingMin and TimeTV are negatively correlated.
That is, the more time students spend reading the less time they spend watching TV. 
  

## Part V
A. Calc correlation coefficient and coefficient of determination.  
Earlier, we looked at correlation between the variables TimeReadingMin and TimeTV and found they were negatively correlated. The coefficient of determination, or R^2, can give us a percentage of variation. That is, we can see how much one variable effects the other. 
```{r echo = FALSE}
cor(ss_edited_df)
```
```{r echo = FALSE}
cor(ss_edited_df)^2
```
  
B. Describe what you conclude about the results.  
The coefficient of determination is a measurement used to explain how
much variability of one factor can be caused by its relationship to another related factor. This correlation, known as the "goodness of fit," is represented as a value between 0.0 and 1.0.

Looking at coeffiecient of determination between TimeTV and Happiness shared variability is about 40% which would imply that TV time variability effects Happiness upto 40% only, while remaining 60% variability in Happiness must be caused by some other variable.  
  
# Part VI
A. Based on analysis, does watching more TV cause students to read less?  
```{r echo = FALSE}
cor(ss_edited_df$TimeReadingMin, ss_edited_df$TimeTV)^2
```
Looking at coefficient of determination (rˆ2) we can say that variability in TimeReading can cause upto 77% variability in TimeTV.
There could be other variables that may cause 23% variability in TimeTV.
  
# Part VII
A. Pick 3 variable and perform a partial correlation.  
I will select TimeReadingMin, TimeTV, and Happiness. Run partial correlation between TimeTV and Happiness while controlling TimeReading 
```{r echo = FALSE}
ss_edited_df2 <- ss_edited_df[,1:3]
```

B. Be sure to document which variable you are "controlling."  
Run partial correlation between TimeTV and Happiness while controlling TimeReading 
```{r echo = FALSE}
pcor(c("TimeTV","Happiness","TimeReadingMin"), var(ss_edited_df2))
pcor(c("TimeTV","Happiness","TimeReadingMin"), var(ss_edited_df2))^2
```  

C. Does this change your interpretation? How, or why not?
If we keep TimeReading controlling , the correlation coefficient between TV time and happiness decrease to 0.59 and coefficient of determination has decreased to 35%. This decrease suggests that variation in Happiness was also effected positively by TimeReading by about 5%.              